### 2015-11-10 paper.Rmd
gof and model selection first hash

---
author:
  - name: David L Miller
    affiliation: Centre for Research into Ecological and Environmental Modelling, University of St Andrews
    address: >
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: dave@ninepointeightone.net
    url: http://converged.yt
title:
  formatted: "Distance Sampling in \\proglang{R}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Distance Sampling in R"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [distance sampling, abundance estimation, line transects, point transects, "\\proglang{R}"]
  plain:     [distance sampling, abundance estimation, line transects, point transects, R]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
output: rticles::jss_article
bibliography: jstatsoft.bib
---

@@

For many years distance sampling analyses have been available via the Windows program Distance (or "DISTANCE"); for clarity henceforth "Distance for Windows" @Thomas:2010cf). From version 5 of Distance for Windows, R packages have been included to perform particular analyses (CITE Distance user manual). This paper shows how to fit detection functions, perform model checking and selection and estimate abundance in \proglang{R} using the package \pkg{Distance2}.

@@

Using this histogram we can crudely estimate the drop-off in detectability by eye by tracing a line that approximates the tops of the histogram bars -- this is the detection function. \pkg{Distance2} estimates the parameters for a fixed-form detection function using maximum likelihood estimation. We address possible models in detail below

@@

We demonstrate \pkg{Distance2} using two data sets: one line transect and one point transect. These data sets have been chosen to be representative of the kind of data seen in practice.

@@

The rest of the paper is structured as follows: we first look into how data needs to be organised to allow it to be used with \pkg{Distance2}; models for the detection function are then covered including their formulation and examples of fitting in \proglang{R}. We then spend some time looking at model checking and goodness of fit testing, as well as model selection. Having illustrated how to obtain a good detection function, we show how to estimate abundance using that model, including using post-stratification. The final two sections of the article look at extensions (both in terms of methodology and software) and put the package in a broader context amongst other software packages used for animal abundance estimation.

@@

The two example data sets used here are distributed with \pkg{Distance2} so readers can reproduce our results easiy. However in general we expect that data will be collected in the field and need to be formatted correctly for use with \pkg{Distance2}. The package allows for a flexible format for data input ranging from very simple to complex:

@@

```{r minke-data-head, message=FALSE}
library(Distance2)
head(minke)
```
and the amakihi data:
```{r amakihi-data-head}
head(amakihi)
```

@@

@@

It is important that the detection function accurately models detectability near zero distance. We are less worried by its behaviour further away from 0. To ensure that the model is not overly influence by distances far from zero we truncate the distances beyond a given distance $w$ (known as the *truncation distance*). Further details on truncation can be found in **CITATION**.

@@

There is a wide literature on possible formulations for the detection function [@Buckland:1992fa; @Eidous:2005bj; @Becker:2009cj; @Giammarino:2014eg; @Miller:2015hw; @Becker:2015fi]. \code{Distance2} includes the most popular of these models, and includes an extendable class system to add new detection functions while avoiding code duplication (see "Extensions").

Here we'll show two popular detection function approaches: "key function plus adjustments" (K+A) and mixture models.

@@

Key function plus an adjustments (or adjustment series) models are formulated by taking a "key" function and optionally adding "adjustments" to it to improve the fit. Mathematically we formulate this as:
$$
g(y; \boldsymbol{\theta}) = k(y; \boldsymbol{\theta}_k)\left( 1+ \alpha_K(y; \boldsymbol{\theta}_\alpha)\right),
$$
where $k$ is the key function and $\alpha_K$ is sum series of functions and the subscripts on the parameter vector indicate those parameters belonging to each part of the model.

Models for $k$ are as follows
$$
k(y) = \left\{
\begin{array}{l l}
  \exp\left(-\frac{y^2}{2 \sigma^2}\right) & \quad \text{half-normal} \\
  1-\exp\left(\left(-\frac{y}{\sigma}\right)^{-b}\right) & \quad \text{hazard-rate} \\
  1/w & \quad \text{uniform}
\end{array} \right.
$$
Possible modelling options for key and adjustments are given in Table \ref{ka-tab} and illustrated in **Figure ???**.

```{r hn-hr-par-comp, fig.width=10, fig.height=6.5, fig.cap="Half-normal (top row) and hazard-rate (bottom row) detection functions without adjustments, varying scale ($\\sigma$) and (for hazard-rate) shape ($b$) parameters (values are given above the plots). On the top row from left to right, the study species becomes more detectable (higher probability of detection at larger distances). The bottom row shows the hazard-rate model's more pronounced shoulder.", echo=FALSE}
par(mfrow=c(2,4), mar=c(3.5, 3, 2, 1) + 0.1)

@@

## hazard-rate
g.hr <- function(x, sigma, b) 1 - exp(-(x/sigma)^-b)
for(this.sig in c(0.1, 0.5)){
  for(this.b in c(5, 1)){
    this.g <- function(x) g.hr(x, sigma=this.sig, b=this.b)
    curve(this.g, from=0, to=1, xlab="", ylab="",
          main=bquote(sigma == .(this.sig) ~ .(", b") == .(this.b)),
          xlim=c(0,1), ylim=c(0,1), asp=1)
    title(xlab="Distance", line=2)
    title(ylab="Detection probability", line=2)
  }
}

@@

When adjustment terms are used it may be necessary to standardise the results to ensure that $g(0)=1$, so we can redefine the detection function as:
$$
g(y; \boldsymbol{\theta}) = \frac{k(y; \boldsymbol{\theta})\left( 1+ \alpha_K(y; \boldsymbol{\theta})\right)}{k(0; \boldsymbol{\theta})\left( 1+ \alpha_K(0; \boldsymbol{\theta})\right)}
$$
A distinct disadvantage of K+A models is that we must resort to constrained optimisation (via the \pkg{Rsolnp} package) in order to ensure that the resulting detection function is monotonic non-increasing over the whole range.

@@

```{r adjust-mix-comp, fig.width=10, fig.height=5, fig.cap="Top row shows possible shapes for the detection function when adjustments are included for half-normal and hazard-rate models. Bottom row shows possible shapes when 2-component mixtures of half-normal functions are used. Notation is as in main text.", echo=FALSE}
par(mfrow=c(2,4), mar=c(3.5, 3, 2, 1) + 0.1)

@@

## hazard-rate with cosine
g <- function(x) (1 - exp(-(x/0.005)^-1.1))*(1-0.05*cos((2*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 1 cosine adjustment",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)
# 2 cosines
g <- function(x) (1 - exp(-(x/0.005)^-1.9))*
                  (1+0.05*cos((2*pi*x)/0.025) + 0.25*cos((3*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 2 cosine adjustments",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)

# mixture models
plotmix <- function(sig1, sig2, mix){

  gmix <- function(x) mix*g.hn(x, sig1) + (1-mix)*g.hn(x,sig2)

  curve(gmix, from=0, to=1, xlab="", ylab="",
        main=bquote(sigma[1] == .(sig1) ~ sigma[2] == .(sig2) ~  phi[1] == .(mix)),
        xlim=c(0,1), ylim=c(0,1), asp=1)
  title(xlab="Distance", line=2)
  title(ylab="Detection probability", line=2)
}
plotmix(0.8, 0.15, 0.3)
plotmix(0.6, 0.1,0.7)
plotmix(10, 0.2, 0.15)
plotmix(0.7, 0.05, 0.6)

### Mixture models

We can also formulate the detection function as a mixture of (say) half-normal functions [@Miller:2015hw]:
$$
g(y; \boldsymbol{\theta}) = \sum_{k=1}^K \varphi_j \exp\left(-\frac{y^2}{2 \sigma_j^2}\right)
$$
where $\sum_{k=1}^K \varphi_j =1$ and $\boldsymbol{\theta} = (\varphi_1,\ldots,\varphi_{K-1},\sigma_1,\ldots,\sigma_K$. Such models do not require constrained optimisation (as K+A models do) as the mixtures are monotonic decreasing by construction.

As with the K+A model, we describe model selection (in terms of number of mixture components) in "Model checking and selection" below.

@@

Covariates can be included in this formulation by considering the scale parameter from the half-normal or hazard-rate detection functions as a linear model (on the exponential scale) of the ($k$) covariates ($\mathbf{z}$):
$$
\sigma(\mathbf{z}) = \exp(\beta_0 + \sum_k \beta_k z_k).
$$
In the next section we'll discuss model selection for these models.

@@

Another important consideration is that K+A models that include covariates and one or more adjustments cannot be guaranteed to be monotonic non-increasing for all covariate combinations, as we don't have any model for the distribution of the covariates. For this reason, we advise against using both adjustments and covariates in a detection function [see @Miller:2015hw for an example of when this can be problematic].

@@

The workhorse of detection function fitting in \pkg{Distance2} is the \code{ds} function. Here we show off the formulations for the detection function that we've seen above for both the minke whale and amakihi data.

@@

We can fit a model to the minke whale data, setting the truncation at 1.5km and using the default options in \code{ds} very simply:
```{r minke-hn}
minke_hn <- ds(minke, truncation=1.5)
```
**Figure???** shows the result of calling \code{plot} on the resulting model object. We can also call \code{summary} on the model object to get summary information about the fitted model (though we postpone this until the next section).

```{r minke-hn-plot, fig.width=4, fig.height=4, fig.cap="Plot of fitted detection function overlayed on the histogram of observed distances for the minke whale data."}
plot(minke_hn)
```

We can specify the form of the detection function via the \code{model=df(...)} argument to \code{ds}. For example, fitting a hazard rate model with cosine adjustments of order 2 can be specified as:
```{r minke-hr-cos}
minke_hrcos <- ds(minke, truncation=1.5, model=df(model=~hr+cos(2)))
```

@@

By default \code{ds} assumes that the data given to it is line transect, but we can switch to points using \code{transect="points"}. Including covariates in the scale is via the `scale=~...` argument to `df`; a hazard-rate model for the amakihi that includes observer as a covariate (and truncating at 82.5m, given in @Marques:2007ey) can be specified by:
```{r amakihi-hr}
amakihi_hr_obs <- ds(amakihi, truncation=82.5, model=df(model=~hr, scale=~obs), transect="point")
```

@@

```{r minke-hn-summary}
summary(minke_hn)
```

* summary results
* qq plots
* hessian/convergence
* monotonicity
* AIC -- what else?

@@

* Horvitz-Thompson
* Refer to Fewster et al 2009

>>>>

---
author:
  - name: David L Miller
    affiliation: Centre for Research into Ecological and Environmental Modelling, University of St Andrews
    address: >
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: dave@ninepointeightone.net
    url: http://converged.yt
title:
  formatted: "Distance Sampling in \\proglang{R}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Distance Sampling in R"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [distance sampling, abundance estimation, line transects, point transects, "\\proglang{R}"]
  plain:     [distance sampling, abundance estimation, line transects, point transects, R]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{longtable}
output: rticles::jss_article
bibliography: jstatsoft.bib
---

@@

For many years distance sampling analyses have been available via the Windows program Distance (or "DISTANCE"); for clarity henceforth "Distance for Windows" @Thomas:2010cf). From version 5 of Distance for Windows, R packages have been included to perform particular analyses (CITE Distance user manual). This paper shows how to fit detection functions, perform model checking and selection and estimate abundance in \proglang{R} using the package \pkg{Distance}. \pkg{Distance} is a wrapper package around the more complex (and more powerful) \pkg{mrds} and offers a subset of the analyses possible with that package.

**Need to include more history here (TRANSECT etc)?**

@@

Using this histogram we can crudely estimate the drop-off in detectability by eye by tracing a line that approximates the tops of the histogram bars -- this is the detection function. \pkg{Distance} estimates the parameters for a fixed-form detection function using maximum likelihood estimation. We address possible models in detail below.

@@

We demonstrate \pkg{Distance} using two data sets: one line transect and one point transect. These data sets have been chosen to be representative of the kind of data seen in practice.

@@

The rest of the paper is structured as follows: we first look into how data needs to be organised to allow it to be used with \pkg{Distance}; models for the detection function are then covered including their formulation and examples of fitting in \proglang{R}. We then spend some time looking at model checking and goodness of fit testing, as well as model selection. Having illustrated how to obtain a good detection function, we show how to estimate abundance using that model, including using post-stratification. The final two sections of the article look at extensions (both in terms of methodology and software) and put the package in a broader context amongst other software packages used for animal abundance estimation.

@@

The two example data sets used here are distributed with \pkg{Distance} so readers can reproduce our results easiy. However in general we expect that data will be collected in the field and need to be formatted correctly for use with \pkg{Distance}. The package allows for a flexible format for data input ranging from very simple to complex:

@@

```{r load-data, echo=FALSE}
# quietly load the data
load("data/minke.RData")
load("data/amakihi.RData")
```
```{r minke-data-head, message=FALSE}
library(Distance)
head(minke)
```
and the amakihi data:
```{r amakihi-data-head}
head(amakihi)
```

@@

**Do we need to talk about binned distances too?**

@@

It is important that the detection function accurately models detectability near zero distance. We are less worried by its behaviour further away from 0. To ensure that the model is not overly influence by distances far from zero we truncate the distances beyond a given distance $w$ (known as the *truncation distance*). Fitting to distances further away from zero does not improve the precision of abundance estimates considerably [@Buckland:2001vm, 103–108, 151–153].

@@

There is a wide literature on possible formulations for the detection function [@Buckland:1992fa; @Eidous:2005bj; @Becker:2009cj; @Giammarino:2014eg; @Miller:2015hw; @Becker:2015fi]. \code{Distance} includes the most popular of these models, and includes an extendable class system to add new detection functions while avoiding code duplication (see "Extensions").

Here we'll detail the most popular detection function approach: "key function plus adjustments" (K+A).

@@

Key function plus adjustment terms (or adjustment series) models are formulated by taking a "key" function and optionally adding "adjustments" to it to improve the fit. Mathematically we formulate this as:
$$
g(y; \boldsymbol{\theta}) = k(y; \boldsymbol{\theta}_k)\left( 1+ \alpha_K(y; \boldsymbol{\theta}_\alpha)\right),
$$
where $k$ is the key function and $\alpha_K$ is sum series of functions and the subscripts on the parameter vector indicate those parameters belonging to each part of the model.

Models for $k$ are as follows
$$
k(y) = \left\{
\begin{array}{l l}
  \exp\left(-\frac{y^2}{2 \sigma^2}\right) & \quad \text{half-normal} \\
  1-\exp\left(\left(-\frac{y}{\sigma}\right)^{-b}\right) & \quad \text{hazard-rate} \\
  1/w & \quad \text{uniform}
\end{array} \right.
$$
Possible modelling options for key and adjustments are given in Table \ref{ka-tab} and illustrated in **Figure ???**. We select the number of adjustment terms ($K$) by AIC (further details in "Model checking and model selection").

```{r hn-hr-par-comp, fig.width=10, fig.height=5, fig.cap="Half-normal (top row) and hazard-rate (bottom row) detection functions without adjustments, varying scale ($\\sigma$) and (for hazard-rate) shape ($b$) parameters (values are given above the plots). On the top row from left to right, the study species becomes more detectable (higher probability of detection at larger distances). The bottom row shows the hazard-rate model's more pronounced shoulder.", echo=FALSE}
par(mfrow=c(2,4), mar=c(3.5, 3, 2, 1) + 0.1)

@@

## hazard-rate
g.hr <- function(x, sigma, b) 1 - exp(-(x/sigma)^-b)
for(this.sig in c(0.1, 0.5)){
  for(this.b in c(5, 1)){
    this.g <- function(x) g.hr(x, sigma=this.sig, b=this.b)
    curve(this.g, from=0, to=1, xlab="", ylab="",
          main=bquote(sigma == .(this.sig) ~ .(", b") == .(this.b)),
          xlim=c(0,1), ylim=c(0,1), asp=1)
    title(xlab="Distance", line=2)
    title(ylab="Detection probability", line=2)
  }
}
```

@@

When adjustment terms are used it may be necessary to standardise the results to ensure that $g(0)=1$, so we can redefine the detection function as:
$$
g(y; \boldsymbol{\theta}) = \frac{k(y; \boldsymbol{\theta})\left( 1+ \alpha_K(y; \boldsymbol{\theta})\right)}{k(0; \boldsymbol{\theta})\left( 1+ \alpha_K(0; \boldsymbol{\theta})\right)}
$$
A disadvantage of K+A models is that we must resort to constrained optimisation (via the \pkg{Rsolnp} package) in order to ensure that the resulting detection function is monotonic non-increasing over the whole range.

@@

```{r adjust-mix-comp, fig.width=9, fig.height=3, fig.cap="Possible shapes for the detection function when adjustments are included for half-normal and hazard-rate models.", echo=FALSE}
par(mfrow=c(1,4), mar=c(3.5, 3, 2, 1) + 0.1)

@@

## hazard-rate with cosine
g <- function(x) (1 - exp(-(x/0.005)^-1.1))*(1-0.05*cos((2*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 1 cosine adjustment",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)
# 2 cosines
g <- function(x) (1 - exp(-(x/0.005)^-1.9))*
                  (1+0.05*cos((2*pi*x)/0.025) + 0.25*cos((3*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 2 cosine adjustments",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)
```

@@

Covariates can be included in this formulation by considering the scale parameter from the half-normal or hazard-rate detection functions as a linear model (on the exponential scale) of the ($J$) covariates ($\mathbf{z}$):
$$
\sigma(\mathbf{z}) = \exp(\beta_0 + \sum_{j=1}^J \beta_j z_j).
$$
In the next section we'll discuss model selection.

@@

Another important consideration is that K+A models which include covariates and one or more adjustments cannot be guaranteed to be monotonic non-increasing for all covariate combinations, as we don't have any model for the distribution of the covariates. For this reason, we advise against using both adjustments and covariates in a detection function [see @Miller:2015hw for an example of when this can be problematic].

@@

The workhorse of detection function fitting in \pkg{Distance} is the \code{ds} function. Here we show off the formulations for the detection function that we've seen above for both the minke whale and amakihi data.

@@

We can fit a model to the minke whale data, setting the truncation at 1.5km and using the default options in \code{ds} very simply:
```{r minke-hn}
minke_hn <- ds(minke, truncation=1.5)
```
Note that \code{ds} will automatically select adjustment terms by AIC and shows it's selection steps as it goes.

**Figure???** shows the result of calling \code{plot} on the resulting model object. We can also call \code{summary} on the model object to get summary information about the fitted model (though we postpone this until the next section).

We can specify the form of the detection function via the \code{key=} argument to \code{ds}. For example, a hazard rate model can be fitted as:
```{r minke-hr-cos}
minke_hrcos <- ds(minke, truncation=1.5, key="hr")
```
Again, \code{ds} fits a model with cosine adjustments (the default) but finds the AIC improvement to be insufficient to select the adjustment.

@@

By default \code{ds} assumes that the data given to it is line transect, but we can switch to points using \code{transect="point"}. Including covariates in the scale is via the \code{formula=~...} argument to \code{ds}; a hazard-rate model for the amakihi that includes observer as a covariate (and truncating at 82.5m, given in @Marques:2007ey) can be specified by:
```{r amakihi-hr-obs}
amakihi_hr_obs <- ds(amakihi, truncation=82.5, transect="point",
                     key="hr", formula=~obs)
```
As with the minke whale model, we can plot the resulting detection function. **Since for the amakihi we used covariates in the detection function, the plot will show the detection function averaged over levels/values of the covariate and separately for each level set (for continuous covariates, the 25\%, 50\% and 75\% quantiles are displayed).**

**It would be nice if I could get "nice" plotting working for `Distance` as they are much easier to interpret (IMO)**

**When multiple covariates are used in the model, a series of plots are produced. We can see this in the bottom row of Figure ???, where we plot the results of fitting following model:**
```{r amakihi-hr-obs-mas, warning=FALSE}
amakihi_hr_obs_mas <- ds(amakihi, truncation=82.5, transect="point",
                         key="hr", formula=~obs+mas)
```

```{r minke-hn-plot, echo=FALSE, warning=FALSE, fig.width=9, fig.height=3, fig.cap="Plot of fitted detection functions overlayed on the histograms of observed distances. Left minke whale data, half-normal model; centre, amakihi data hazard-rate with observer as a covariate; right, amakihi data, hazard-rate model with observer and minutes after sunrise as covariates"}
par(mfrow=c(1,3))
plot(minke_hn)
plot(amakihi_hr_obs)
plot(amakihi_hr_obs_mas)
```

@@

As with models fitted using \code{lm} or \code{glm} in \proglang{R} we can use \code{summary} to give us some useful information about our fitted model. For example for our hazard-rate model for the amakihi, with observer as a covariate:
```{r amakihi-summary}
summary(amakihi_hr_obs)
```
This summary information includes various details of the data and model specification, as well as the values of the coefficents ($\beta_j$) and their uncertainties, an "average" value for the detectability (see "Estimating abundance and variance" for details on how this is calculated) and it's uncertainty. The final line gives an estimate of abundance for the area covered by the survey (again, this will be covered in detail in the next section).

## Goodness of fit

To judge goodness of fit for detection functions when exact distances are used, we want to compare the cumulative distribution function (CDF) and empirical distribution function (EDF) for the detection function via a quantile-quantile plot (Q-Q plot). In our case the CDF evaluates the probability of observing an animal at a distance less than or equal to some given value. The EDF gives the proportion of observations for which the CDF is less than or equal to that of a given distance. In essence we're asking "is the number of observations up to a given distance in line with what the model says they should be?" (where “given values” are the observed distances). As usual for Q-Q plots, "good" models will have values close to the line $y=x$, poor models will show more deviations from that line.

We can inspect Q-Q plots visually, though for a large number of models this can be tiresome and prone to subjective judgements. Instead we can quantify the Q-Q plot's information using a Kolmogorov-Smirnov or Cramer-von Mises test [@Burnham:2004vd]. Both test to see if the points from the EDF and CDF are from the same distribution. The Kolmogorov-Smirnov uses the test statistic of the largest difference between a point on the Q-Q plot and the line $y=x$, whereas the Cramer-von Mises test uses the sum of all the differences. As it takes into account more information and is therefore more powerful, the Cramer-von Mises is generally preferred. A significant result from either tests indicates that the EDF and CDF do not come from the same distribution (and therefore the model is does not fit the data well).

We can generate a Q-Q plot and test results using the \code{gof_ddf} function. We can see the goodness of fit tests below for two models for the amakihi data, first fitting a half-normal model without covariates or adjustments (not setting \code{adjustment=NULL} will force \code{ds} to fit a model with no adjustments), then calculating goodness of fit for that model and our hazard-rate model with observer and minutes after sunrise included:
```{r amakihi-gof, fig.keep="none"}
amakihi_hr <- ds(amakihi, truncation=82.5, transect="point", key="hn", adjustment=NULL)
gof_ds(amakihi_hr)
gof_ds(amakihi_hr_obs_mas)
```
The corresponding Q-Q plots are shown in **Figure ???**.

```{r, amakihi-qq-comp, fig.width=6, fig.height=3, fig.cap="Comparison of quantile-quantile plots for a half-normal model (no adjustments, no covariates) and hazard-rate model with observer and minutes after sunrise for the amakihi data.", echo=FALSE, results="hide"}
par(mfrow=c(1,2))
gof_ds(amakihi_hr, asp=1)
gof_ds(amakihi_hr_obs_mas, asp=1)
```

**Do we need to add something about binned distances and Q-Q plots here?**

## Model selection

Once we have a set of models which fit well, we can use Akaike's Information Criterion (AIC) to select between models. \pkg{Distance} includes a function to create table of summary information for fitted models, making it easy to get an overview of a large number of models at once. The \code{summarize_ds_models} function takes models as input and can be especially useful when paired with \pkg{knitr}'s \code{kable} function to create summary tables for publication. An example of this output is shown in **Table ???**.

```{r results="asis", echo=FALSE}
kable(summarize_ds_models(amakihi_hr, amakihi_hr_obs, amakihi_hr_obs_mas),
      digits = 3, format = "latex", row.names = FALSE, escape=FALSE,
      caption="Summary for the detection function models fitted to the amakihi data. \"C-vM\" stands for Cramer-von Mises. Models are sorted according to AIC.")
```

* monotonicity -- do we need this?

@@

* Horvitz-Thompson
* Refer to Fewster et al 2009
* calculation of $P_a$
